# Introduction {#intro}

## Terms

"A network is not just a metaphor: it is a precise, mathematical construct of nodes (vertices, actors) N and edges (ties, relations) E that can be directed or undirected." [@jasneyIntroductionSocialNetwork2018]

```{r node-graph, fig.show="hold", echo=FALSE, out.width="50%", fig.cap="a node"}
library(DiagrammeR)
library(grid)
library(gridExtra)
#graph 1 - unconnected
n <-  12
ndf <-
  create_node_df(
    n = n,
    label = paste0("N", n)
  )
graph1 <- create_graph(nodes_df = ndf)
graph1 %>% render_graph()
#graph 2 - connected
edf <-
  create_edge_df(
    from = 1:n,
    to   = n:1,
    rel  = ""
  )
graph2 <- create_graph(nodes_df = ndf, edges_df = edf)
graph2 %>% render_graph(layout = "circle",
                        output = "visNetwork")
```

```{r}
df <- data.frame(nodes = c(1, 2), y = c(1, 1))
library(ggplot2)
p <- ggplot()
p <- p + geom_point(aes(1, 1), 
                    size = 25,
                    shape = 1,
                    label = "N1"
                    )
p <- p + annotate("text", x = 1, y = 1, label = "N1")

p <- p + theme_void()
p
```



## Input formats

Building a network graph is challenging because of the number of potential input types. Users must first distinguish whether the input is an adjacency matrix, incidence matrix, or edge list.

### Adjacency matrix

### Incidence matrix

### Edge list

### Dataframe distinguished

## R packages

Three `R` packages are dedicated to network analysis and are available for download from CRAN:  `igraph`, `ggraph`, and `networkD3`.


You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].
