# igraph {#igraph}

```{r load-igraph-library, include=FALSE}
x <- c("igraph", "tibble", "stringr", "kableExtra", "RColorBrewer", "viridis")
lapply(x, require, character.only = T, quietly = T)
```
`igraph` contains `r length(lsf.str("package:igraph"))` functions. Three different plotting mechanisms exist: `igraph` for two dimensional plots, `tkplot` for interactive 2-dimensional plotting and `rglplot` which is for 3-dimensional plotting and is labeled as "experimental." The `igraph` functions for two dimensional plots is the focus for the next * chapters. 

## Examples

### Simple Graphs

The simple graph below was created from manual entry using igraph's `graph_from_literal` function.

```{r simple-igraph-node, fig.cap="Graph (a) is two nodes connected by edge, graph (b) is two unconnected nodes, graph (c) is two nodes with a directional edge, graph (d) is two nodes with the directional edge reversed, graph (e) is two nodes with a bidirectional edge, and graph (f) is two nodes connected by a weighted edge.", echo=FALSE, fig.align="center"}

library(grid)
library(gridExtra)
filename <- "./imgs/introductory_plots.jpg"
par(mfrow=c(2, 3))
igraph_options(vertex.size = 150, arrow.size = .1, vertex.label.font = 16)
g1 <- graph_from_literal(A--B)
g2 <- graph_from_literal(A, B)
g3 <- graph_from_literal(A-+B)
g4 <- graph_from_literal(A+-B)
g5 <- graph_from_literal(A++B)
g6 <- graph_from_literal(B--A)
plot.igraph(g1, sub = "(a)")
plot.igraph(g2, sub = "(b)")
plot.igraph(g3, sub = "(c)", edge.arrow.size = .5)
plot.igraph(g4, sub = "(d)", edge.arrow.size = .5)
plot.igraph(g5, sub = "(e)", edge.width = .5, edge.arrow.size = .5)
plot.igraph(g6, sub = "(f)", edge.width = 10)

```

### Notable Graphs

```{r notable-graphs-example, results="hide", echo=F, fig.height=6, fig.width = 6, fig.align="center", fig.cap="Nine of 31 'Notable Graphs' from 'make_graph()`"}
par(mfrow = c(3,3), mar = c(1, 1, 1, 1))
x <- c("Bull", "Chvatal", "Coxeter", "Cubical", "Diamond", "Dodecahedral", "Folkman", "Franklin", "Frucht", "Grotzsch", "Heawood", "Herschel", "House", "HouseX", "Icosahedral", "Krackhardt kite", "Levi", "McGee", "Meredith", "Noperfectmatching", "Nonline", "Octahedral", "Petersen", "Robertson", "Smallestcyclicgroup", "Tetrahedral", "Thomassen", "Tutte", "Uniquely3colorable", "Walther", "Zachary")
notable.graphs <- sample(x, size = 9, replace = F)
sapply(notable.graphs, function(x) plot.igraph(make_graph(x), main = x, vertex.size = 10))
```


### Complicated Graph

The complicated graph below was created from the Les Miserable resident dataset in `igraph`. [@knuthStanfordGraphBasePlatform1993] The data contains 77 nodes and 254 edges.


```{r complicated-igraph-node, fig.cap="Network graph of the coappearnce of characters in Les Miserable.", echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}
library(qgraph)
#igraph may have lesmis as resident dataset
file <- "./data/gephy/lesmiserables.gml"
g1 <- read_graph(file = file, format = "gml")
igraph_options(vertex.size = 10)
layout <- layout_with_fr(g1)
e <- get.edgelist(g1)
#l <- qgraph.layout.fruchtermanreingold(e,vcount=vcount(g1))
l <- qgraph.layout.fruchtermanreingold(e,vcount=vcount(g1), area=8*(vcount(g1)^2),repulse.rad=(vcount(g1)^3.1))
#that worked! See https://stackoverflow.com/questions/39290909/igraph-resolving-tight-overlapping-nodes
plot.igraph(g1, 
            asp = 0,
            vertex.label.cex = .35, 
            layout = l,
            axes = F)
```

## Layout of Nodes

Nodes may be laid out in a variety of ways and many algorithms are available, depending on the data and the analyst's personal preference. "It is important to realize that the information in graph layouts is contained in the pattern of which nodes are connected to which others.  The locations . . . are chosen based on essentially aesthetic criteria." [@borgattiAnalyzingSocialNetworks2018]

### Consistent Plotting

Many of the `igraph` layouts are generated through an algorithm and the coordinates change each time it is plotted.  To keep the appearance the same, the `set.seed()` function must be set.

```{r igraph-house-setseed, eval=F}
set.seed(1234)
house <- make_graph("house")
plot.igraph(house)
```

A second method to retain consistency in plotting is to retain the coordinates in a separate variable as opposed to the `set.seed()` strategy.

```{r igraph-coordinates-variable}
data("UKfaculty")
coords <- layout_with_fr(UKfaculty, niter = 1000)

par(mfrow=c(1, 2))
par(oma=c(0,0,2,0))

plot.igraph(UKfaculty, layout = coords, main = "layout = coords #1", vertex.color = "blue")

plot.igraph(UKfaculty, layout = coords, main = "layout = coords #2", vertex.color = "red")

```


### Available Layouts

According to the igraph documentation, 21 available layouts are available and include:

```{r layout-options-igraph, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", fig.cap="Available layouts for nodes in igraph."}

library(stringr)
a <- str_trim(unlist(str_split("add_layout_, component_wise, layout_as_star, layout_as_tree, layout_in_circle, layout_nicely, layout_on_grid, layout_on_sphere, layout_randomly, layout_with_dh, layout_with_fr, layout_with_gem, layout_with_graphopt, layout_with_kk, layout_with_lgl, layout_with_mds, layout_with_sugiyama, layout_, merge_coords, norm_coords, normalize", ",")), side = "both")
m <- matrix(a, nrow = 7, ncol = 3, byrow = T)
m %>%
  kbl(caption = "Available layouts for nodes in `igraph`.") %>%
  kable_styling()
```

Plots were randomly generated from an adjacency matrix as illustrated by [R-graph-gallery](https://www.r-graph-gallery.com/network.html). The `layout_nicely` function is the default.  (verify).  The function 
uses the `layout_with_fr` with graphs less than 1,000 vertices and the `layout_with_drl` for graphs greater than 1,000 vertices.

```{r igraph-mosaic-layout, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", fig.cap="Examples of available layouts."}
# Create data
data <- matrix(sample(0:1, 100, replace=TRUE, prob=c(0.5,0.5)), nrow=10)
network <- graph_from_adjacency_matrix(data , mode='undirected', diag=F )
 
# When ploting, we can use different layouts:
par(mfrow=c(3,3), mar=c(1,1,1,1))
plot(network, layout=layout_nicely, main = "nicely")
plot(network, layout=layout_on_grid, main = "grid")
plot(network, layout=layout_with_drl, main="DrL")
plot(network, layout=layout.circle, main="circle")
plot(network, layout=layout.random, main="random")
plot(network, layout=layout.fruchterman.reingold, main="fruchterman-reingold")
plot(network, layout=layout_as_tree, main="tree")
plot(network, layout=layout_with_kk, main="kamada-kawai")
plot(network, layout=layout_with_lgl, main="large.graph.layout")
```

In plotting graphs with many nodes, the chosen layout will result in overlapping nodes.  The challenge is in spreading the nodes out to the space available.  One strategy was described in a stackoverflow [question](https://stackoverflow.com/questions/39290909/igraph-resolving-tight-overlapping-nodes) using the `qgraph` library.  The code was was used to reproduce the plot below.


```{r spread-node-layout, echo=F, fig.align="center", fig.width = 7, fig.height = 4, fig.cap="Node dispersion of the Fruchterman-Reingold (F-R) layout."}
library(igraph)
library(qgraph)
g <- barabasi.game(355, directed=FALSE)
par(mfrow=c(1, 3), mar = c(1, 1, 1, 1))
plot(g,layout=layout_with_fr,
     vertex.size=4,
     vertex.label=NA,
     main = "layout_with_fr")
e <- get.edgelist(g)
l <- qgraph.layout.fruchtermanreingold(e,vcount=vcount(g))
plot(g,
     layout = l,
     vertex.size = 4,
     vertex.label = NA,
     main ="F-R default"
     )
l <- qgraph.layout.fruchtermanreingold(e,vcount=vcount(g),
      area=8*(vcount(g)^2),repulse.rad=(vcount(g)^3.1))
plot(g,
     layout = l,
     vertex.size = 4, 
     vertex.label = NA,
     main = "F-R modified")
```

## Attributes

Any chart is comprised of geoms.  `igraph` allows for the color, size, shape to be set in of two ways.  The first method is to supply named arguments to the plotting command, in the this case `plot.igraph`. The second method is to set the attributes via assignment to the vertex, edge or graph.  According to the manual, assignments using the second method do not require "vertex" or "edge" to be prepended.

Graph attributes are listed anytime the graph object is sent to the console.  See section *. Othewise, attributes of the graph, vertices, and edges can be determed through three functions:  (1) `list.graph.attributes`, (2) `list.vertex.attributes`, and (3) `list.edge.attributes`.  The function length can be cumbersome and the following function may be helpful in list all of them at once. [@knappNetworkVisualizationCookbook]

```{r igraph-list-attribs-function, eval = F}
#
get_igraph_attrs <- function(igraph){
  if(!is_igraph(igraph)) stop("Not a graph object")
  list(graph_attrs = list.graph.attributes(igraph),
       vertex_attrs = list.vertex.attributes(igraph),
       edge_attrs = list.edge.attributes(igraph))
}

get_igraph_attrs(UKfaculty)
```


```{r example-attribute-setting, eval=F, attr.source = ".numberLines"}
#First Method -- supply named arguments to plot
plot.igraph(make_graph("Bull"),
            vertex.shape = "square"
            )
#Second Method -- set vertex attributes
g <- make_graph("Bull") %>%
        set_vertex_attr(name = "shape", 
                        value = "square")
```


### Vertices

```{r vertex-attributes, echo=F, fig.align="center", fig.cap="Selected node attributes."}
par(mfrow=c(2,2), mar=c(2, 2, 2, 2))
# shapes
shapes <- shapes()
g1 <- make_ring(length(shapes()))
plot.igraph(g1,
            vertex.size = 50,
            vertex.shape = shapes,
            main = "shapes"
            )

# size
sizes <- seq(5, 50, length.out = 10)
g2 <- make_ring(length(sizes))
plot.igraph(g2, 
            vertex.size = sizes,
            vertex.shape = "circle",
            main = "sizes")
# color
colors <- c(NA, 1:8)
g3 <- make_ring(9)
plot.igraph(g3, 
            vertex.color = colors,
            vertex.size = 50,
            vertex.label = paste0(colors),
            layout = layout_nicely,
            main = "colors")
# labels
labels <- state.abb[1:10]
size <- seq(.1, 2, length.out = 10)
dist <- seq(.1, 3, length.out = 10)
g4 <- make_ring(10)
plot.igraph(g4, 
            vertex.label = labels,
            vertex.label.cex = size,
            vertex.label.dist = dist,
            layout = layout_nicely,
            main = "labels"
            
            )

```

```{r more-color-examples, fig.height=3.5, echo=F, fig.align="center", fig.cap="Additional node color palettes."}
par(mfrow=c(1,3), mar=c(2, 2, 3, 2))
#graph
colors <- brewer.pal(name = "Set1", n = 9)
g1 <- make_ring(9)
plot.igraph(g1, 
            vertex.color = colors,
            vertex.shape = "circle",
            vertex.label.cex = .75,
            vertex.size = 30,
            vertex.label = paste0(colors),
            vertex.label.dist = 5,
            layout = layout_nicely,
            main = "brewer\n Set1"
            )
#graph
colors <- viridis(10, option = "D")
g1 <- make_ring(length(colors), circular = T)
plot.igraph(g1, 
            vertex.color = colors,
            vertex.shape = "csquare",
            vertex.label.cex = .75,
            vertex.size = 25,
            vertex.label = paste0(colors),
            vertex.label.dist = 4,
            layout = layout_nicely,
            main = "viridis \n default"
            )
#graph
colors <- rainbow(10)
g3 <- make_ring(length(colors))
plot.igraph(g3, 
            vertex.color = colors,
            vertex.shape = "csquare",
            vertex.label.cex = .75,
            vertex.size = 25,
            vertex.label = paste0(colors),
            vertex.label.dist = 4,
            layout = layout_nicely,
            main = "base \n rainbow"
            )

```


#### Shapes

#### Sizes

#### Colors

The argument to set the vertex color in the `plot.igraph` function is `vertex.color`.  Like plots in base `R`, color may be set by a number (integer) or name (character) with the name being preferred. The default palette can be retrieved through `?igraph::categorical_pal(n)`.  The colors are visible to those who are color-blind and is eight colors.



For plot.igraph and integer values, the default igraph palette is used (see the ‘palette’ parameter below. Note that this is different from the R palette.

If you don't want (some) vertices to have any color, supply NA as the color name.

The default value is “SkyBlue2”.



### Edges

```{r edge-attributes, echo = F, fig.align="center", fig.cap="Selected edge attributes."}
par(mfrow=c(2,2), mar=c(2, 2, 2, 2))
# lty

edge.lty <- 1:6
edf <- tibble(from = rep("+", length(edge.lty)),
                  to = LETTERS[1:length(edge.lty)]
               )
g1 <- graph_from_data_frame(edf)
plot.igraph(g1,
            vertex.size = 25,
            edge.lty = edge.lty,
            edge.label = 1:6,
            arrow.mode = "-",
            layout = layout_as_tree,
            main = "type"
            )
#plot
width <- seq(from = .5, to = 3, length.out=6)
edf <- tibble(from = rep("+", length(width)),
                  to = LETTERS[2:sum(length(width), 1)]
               )
g2 <- graph_from_data_frame(edf)
plot.igraph(g2,
            vertex.size = 25,
            edge.lty = 1,
            edge.label = 1:length(width),
            arrow.mode = "-",
            edge.width = width,
            layout = layout_as_tree,
            main = "width"
            )
#plot
colors <- c(1:8)
edf <- tibble(from = rep("+", length(colors)),
                  to = LETTERS[2:sum(length(colors), 1)]
               )
g3 <- graph_from_data_frame(edf)
plot.igraph(g3,
            vertex.size = 25,
            edge.lty = 1,
            edge.label = 1:length(colors),
            arrow.mode = "-",
            edge.width = 1.5,
            layout = layout_as_tree,
            edge.color = colors,
            main = "colors"
            )
#plot labels
colors <- 1:8
edge.label.cex = seq(.5, 4, by = .5)
g4 <- make_ring(8, circular = TRUE)
edf <- tibble(from = rep("+", length(edge.label.cex)),
                  to = LETTERS[2:sum(length(edge.label.cex), 1)]
               )
#g4 <- graph_from_data_frame(g4)
plot.igraph(g4,
            vertex.size = 15,
            vertex.label = NA,
            edge.lty = 1,
            edge.label = colors,
            arrow.mode = NA,
            edge.width = 3,
            edge.color = colors,
            edge.label.cex = edge.label.cex,
            curved = TRUE,
            main = "labels"
            )



```

