# Glossary {#glossary}

**Attributes** are a characteristic of the node or edge that is often designated by color, size, or shape of the object. For example, a dashed line for an edge or a blue circle for a node.

**Nodes** are also actors or vertices.  They are things like people, places, ideas that have some connection.  In mathmatical formulas, they often take the variable $N$.

**Degree** the number of nodes adjacent to the node under evaluation or the number of lines incident to it.

**Diameter** is the maximum number of edges.

**Edges** are ties, relations or connections between nodes.  Edges are often designated as $E$.

**Indegree** is the number of received ties.

**Networks**

**Network objects** is a class of objects in R designed specifically for network analysis.  They store an adjacency matrix or an edgelist as well as metadata. 

**Outdegree** the number of sent ties.

**Directed** networks 

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].
