---
title: "Network Analysis"
author: "Robert Wiederstein"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
description: "This is a series of tutorials on network applications."
---

# Hello World

Placeholder


## R Markdown
## Including Plots

<!--chapter:end:index.Rmd-->

# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

<!--chapter:end:01-intro.Rmd-->

# Literature

Here is a review of existing methods.

<!--chapter:end:02-literature.Rmd-->

# First Graphs

## Stop light

```{r}
library(sigmajs)
edges <- tibble(id = rep("1", 3),
                    source = rep("1", 3),
                    target = as.character(c(2:4))
                    )
nodes <- tibble(id = as.character(1:4),
                    label = c("light", "red", "yellow", "green"),
                    time = c(100, 30, 10, 20)
)
sigmajs() %>%
  sg_nodes(nodes, id, label, time) %>%
  sg_edges(edges, id, source, target) 

```


## Les Miserabe


<!--chapter:end:05-first-charts.Rmd-->

